---
title: "Untitled"
author: ""
date: "March 7, 2017"
output: html_document
---




# Project Goals

The goal of this project is to work through building functions around the Jackkinfe. We will start with some simple functions and then determine a pattern to build a more general function in the end. 


## Jackknife of the Mean!

- Have an estimator $\hat{\theta}$ of parameter $\theta$  
  want the standard error of our estimate, $se_{\hat{\theta}}$
- The jackknife approximation:
    + omit case $i$, get estimate $\hat{\theta}_{(-i)}$
    + Take the variance of all the $\hat{\theta}_{(-i)}$
    + multiply that variance by $\frac{(n-1)^2}{n}$ to get $\approx$ variance of $\hat{\theta}$
- then $se_{\hat{\theta}}=$ square root of that variance

**PHP 2560 Only** (Why $(n-1)^2 / n$?  Think about just getting the standard
    error of the mean)
    

Write a function called `mean.jackknife` that takes argument `a_vector` and returns a jackknife estimate of the standard error.


```{r, eval=TRUE}
my_mean=c()
mean.jackknife <- function(a_vector) {
  for (i in seq_along(a_vector)){
    my_mean[i]<-mean(a_vector[-i], na.rm=TRUE)
    var.jack<-var(my_mean, na.rm=TRUE)
    n<-length(a_vector)
    rescale.var.jack<-((n-1)^2)/n*var.jack
    jackknife.stderr<-sqrt(rescale.var.jack)
  }
  return(jackknife.stderr)
}




```



#### Jackknife for the mean Example: test your code here


```{r, eval = TRUE}
some_normals <- rnorm(100,mean=7,sd=5)
mean(some_normals)
(formula_se_of_mean <- sd(some_normals)/sqrt(length(some_normals)))
all.equal(formula_se_of_mean,mean.jackknife(some_normals))
```



##Jackknife for Gamma Parameters


The following function is a way to calculate the method of moments estimators for the gamma distribution:

```{r, eval=TRUE}
gamma.est <- function(the_data) {
  m <- mean(the_data)
  v <- var(the_data)
  a <- m^2/v
  s <- v/m
  return(c(a=a,s=s))
}
```


###Jackknife for Gamma Parameters Function

Write a function called `gamma.jackknife` that takes argument `a_vector` and returns jackknife standard error estimates on the gamma parameters. 

```{r, eval=TRUE }
a_gamma<-c()
s_gamma<-c()

gamma.jackknife <- function(a_vector) {
  for (i in seq_along(a_vector)){
   n=length(a_vector)
   a_gamma[i]<-gamma.est(a_vector[-i])["a"]
   s_gamma[i]<-gamma.est(a_vector[-i])["s"]
   var.a<-var(a_gamma, na.rm=TRUE)
   stderr.a<-((n-1)^2/n*var.a)^(1/2)
   var.s<-var(s_gamma, na.rm=TRUE)
   stderr.s<-((n-1)^2/n*var.s)^(1/2)
   jackknife.stderr<-c(stderr.a, stderr.s)
   
  }
   return(c(jackknife.stderr))
}


```



###Jackknife for Gamma Parameters Example


```{r, eval=TRUE}
input <- rgamma(1000, shape=0.1, scale=10)
gamma.est(input)
gamma.jackknife(input)
all.equal(gamma.est(input),gamma.jackknife(input))

```



##Jackknife for linear regression coefficients

Write a function called `jackknife.lm` that takes arguments `df`, `formula` and `p` and returns jackknife standard error estimates on the coefficients of a linear regression model. 

```{r, eval=TRUE}

jackknife.lm <- function(df,formula,p) {
  n<-nrow(df)
  coeff<-matrix(NA, nrow=n, ncol=p)
  
  for (i in 1:nrow(df)){
   model<-lm(as.formula(formula), df[-i,])
   coeff[i,]<-model$coefficients}
  
  var.coeff<-apply(coeff, 2, var) # 2 indicates go through the column and calculate the variance (to go through the rows use 1)
  jackknife.stderr<-((n-1)^2/n*var.coeff)^(1/2)
  return(jackknife.stderr)
  }

```



###Jackknife for linear regression coefficients Example


```{r, eval=TRUE}
output <- 1.2 + 0.6*input +  rnorm(1000, 0, 2.1)
data <- data.frame(output,input)
my.lm <- lm(output~input, data=data)
coefficients(my.lm)
# "Official" standard errors
sqrt(diag(vcov(my.lm)))
jackknife.lm(df=data,formula="output~input",p=2)


```




##Refactoring the Jackknife


- Omitting one point or row is a common sub-task

- The general pattern:

```
figure out the size of the data
for each case
   omit that case
   repeat some estimation and get a vector of numbers
take variances across cases
scale up variances
take the square roots
```

- Refactor by extracting the common "omit one" operation

- Refactor by defining a general "jackknife" operation



###The Common Operation


- _Problem_: Omit one particular data point from a larger structure

- _Difficulty_: Do we need a comma in the index or not?

- _Solution_: Works for vectors, lists, 1D and 2D arrays, matrices, data frames:



## Goal: 


- Make the function select the correct dimensions
    - length for a 1d object
    - number of rows for 2d
- Write a function `omit.case` that omits a point given the data and returns the data minus that point. Make sure it can handle higher dimensions. 
    ```{r, eval=TRUE}
          omit.case <- function(the_data,omitted_point) {
                  if (is.null(dim(the_data)) | (length(dim(the_data)) == 1)) {
                    output<-the_data[-omitted_point]
                  } else {output<-the_data[-omitted_point,]}
            return (output)
          }
   
    ```
- Write a function `omit_and_est` that takes the data with an omitted point and returns whatever function your estimator does.  
    ```{r,eval=TRUE}

jackknife <- function(estimator,the_data) {  # Define n to be lenth or number of rows here
  if (is.null(dim(the_data))) {n<- length(the_data)}
  else { n<-nrow(the_data)}
  omit_and_est <- function(omit) {# This function should take the output of omit.case and use it as input for the estimator 
  estimator(omit.case(the_data, omit))}
  jackknife.ests <- matrix(sapply(1:n, omit_and_est), ncol=n)
  var.of.reestimates <- apply(jackknife.ests,1,var)
  jackknife.var <- ((n-1)^2/n)* var.of.reestimates
  jackknife.stderr <- sqrt(jackknife.var)
  return(jackknife.stderr)
}
```



---

###It works


```{r, eval=TRUE}
jackknife(estimator=mean,the_data=some_normals)
all.equal(jackknife(estimator=mean,the_data=some_normals),
          mean.jackknife(some_normals))
```

---



```{r, eval=TRUE}
all.equal(jackknife(estimator=gamma.est,the_data=data$input),
          gamma.jackknife(data$input))
```

---


```{r, eval=TRUE}
all.equal(jackknife(estimator=gamma.est,the_data=data$input),
          gamma.jackknife(data$input), check.names=FALSE)
```


```{r, eval=TRUE}
est.coefs <- function(the_data) {
  return(lm(output~input,data=the_data)$coefficients)
}
est.coefs(data)
all.equal(est.coefs(data), coefficients(my.lm))
```

---


```{r, eval=TRUE}
jackknife(estimator=est.coefs,the_data=data)
all.equal(jackknife(estimator=est.coefs,the_data=data),
          jackknife.lm(df=data,formula="output~input",p=2))
```


##Further Refactoring of jackknife()


The code for `jackknife()` is still a bit clunky:
- Ugly `if-else` for finding `n`
- Bit at the end for scaling variances down to standard errors

- write a function that calculates the `n` needed for the above code:

    ```{r, eval=TRUE}
      data_size <- function(the_data) {
        if (is.null(dim(the_data))) {n<- length(the_data)}
        else { n<-nrow(the_data)}
        return(n)
      }
    ```
- Write a function that calculate the variance of all the estimates and returns the standard error
```{r, eval=TRUE}
scale_and_sqrt_vars <- function(jackknife.ests,n) {
  var.of.reestimates <- apply(jackknife.ests,1,var)
  jackknife.var <- ((n-1)^2/n)* var.of.reestimates
  jackknife.stderr <- sqrt(jackknife.var)
  return(jackknife.stderr)
 }
```



###Now invoke those functions

```{r, eval=TRUE}
jackknife <- function(estimator,the_data) {
  n <- data_size(the_data)
  omit_and_est <- function(omit) {
    estimator(omit.case(the_data,omit))
  }
  jackknife.ests <- matrix(sapply(1:n, omit_and_est), ncol=n)
  return(scale_and_sqrt_vars(jackknife.ests,n))
}
```

